---
title: "Report"
output: html_document
---

```{r setup, include=FALSE}
here::i_am("Rmd/report.Rmd")
knitr::opts_chunk$set(echo = TRUE)
```

```{r load data, include=FALSE}
# Load Data
  # dates data
todays_date <- base::readRDS(here::here("raw_data", "dates_list.rds"))[[1]]
us_close_date <- base::readRDS(here::here("raw_data", "dates_list.rds"))[[2]]
or_close_date <- base::readRDS(here::here("raw_data", "dates_list.rds"))[[3]]
us_data_date <- base::readRDS(here::here("raw_data", "dates_list.rds"))[[4]]
or_data_date <- base::readRDS(here::here("raw_data", "dates_list.rds"))[[5]]
  # diff data 
us_date_diff <- todays_date - us_close_date
or_date_diff <- todays_date - or_close_date
  # dates data
date_today <- base::readRDS(here::here("raw_data", "dates_list.rds"))[[1]]
us_data_date <- base::readRDS(here::here("raw_data", "dates_list.rds"))[[4]]
or_data_date <- base::readRDS(here::here("raw_data", "dates_list.rds"))[[5]]
  # current data 
us_current <- base::readRDS(here::here("clean_data", "clean_data_list.rds"))[[3]]
or_current <- base::readRDS(here::here("clean_data", "clean_data_list.rds"))[[4]]
  # new data
or_new <- base::readRDS(here::here("clean_data", "clean_data_list.rds"))[[6]]
  # list data 
us_list <- readRDS(here::here("clean_data","us_list.rds"))
or_list <- readRDS(here::here("clean_data", "or_list.rds"))
```

The Coronavirus is an infectious disease^[**World Health Organization**  [_Coronavirus disease (COVID-19)_](https://www.who.int/health-topics/coronavirus#tab=tab_1). 2022.], that shut down the United States on `r us_close_date` (`r us_date_diff` days ago) and Oregon on `r or_close_date`(`r or_date_diff` days ago). While most of the state and country have reopened, we still hear about spikes in cases, and warnings about traveling. This study uses two data sets from the NY Times github repository  ^[[New York Times Covid 19 Data](https://github.com/nytimes/covid-19-data).], census data from the RStudio package tidycensus^[[tidycensus](https://walker-data.com/tidycensus/).], and interactive data visualizations to further investigate where and when the coronavirus is impacting Americans and Oregonians. 

---

<center>

Project Updated: `r todays_date` $\quad$    |   $\quad$ US Data Updated: `r us_data_date`   $\quad$|$\quad$   County Data Update : `r or_data_date`

</center>

---

# United States

## Populated Places

#### What states have the **most** covid cases? What states have the **most** covid related deaths? What states have the **most** people fully and partially vaccinated?

Since `r us_data_date`: 

The state with the most reported covid cases is `r us_list[[1]][1,1]` (<span style="color: turquoise ;">`r us_list[[1]][1,2]`</span>), followed by `r us_list[[1]][2,1]` (<span style="color: turquoise ;">`r us_list[[1]][2,2]`</span>), `r us_list[[1]][3,1]` (<span style="color: turquoise ;">`r us_list[[1]][3,2]`</span>), `r us_list[[1]][4,1]` (<span style="color: turquoise ;">`r us_list[[1]][4,2]`</span>), and then `r us_list[[1]][5,1]` (<span style="color: turquoise ;">`r us_list[[1]][5,2]`</span>).

The state with the most reported covid related deaths is `r us_list[[3]][1,1]` (<span style="color: turquoise ;">`r us_list[[3]][1,2]`</span>), followed by `r us_list[[3]][2,1]` (<span style="color: turquoise ;">`r us_list[[3]][2,2]`</span>), `r us_list[[3]][3,1]` (<span style="color: turquoise ;">`r us_list[[3]][3,2]`</span>), `r us_list[[3]][4,1]` (<span style="color: turquoise ;">`r us_list[[3]][4,2]`</span>), and then `r us_list[[3]][5,1]` (<span style="color: turquoise ;">`r us_list[[3]][5,2]`</span>). 

The state with the most people fully vaccinated is `r us_list[[5]][1,1]` (<span style="color: turquoise ;">`r us_list[[5]][1,2]`</span>), followed by `r us_list[[5]][2,1]` (<span style="color: turquoise ;">`r us_list[[5]][2,2]`</span>), `r us_list[[5]][3,1]` (<span style="color: turquoise ;">`r us_list[[5]][3,2]`</span>), `r us_list[[5]][4,1]` (<span style="color: turquoise ;">`r us_list[[5]][4,2]`</span>), and then `r us_list[[5]][5,1]` (<span style="color: turquoise ;">`r us_list[[5]][5,2]`</span>). 

The state with the most people partially vaccinated is `r us_list[[7]][1,1]` (<span style="color: turquoise ;">`r us_list[[7]][1,2]`</span>), followed by `r us_list[[7]][2,1]` (<span style="color: turquoise ;">`r us_list[[7]][2,2]`</span>), `r us_list[[7]][3,1]` (<span style="color: turquoise ;">`r us_list[[7]][3,2]`</span>), `r us_list[[7]][4,1]` (<span style="color: turquoise ;">`r us_list[[7]][4,2]`</span>), and then `r us_list[[7]][5,1]` (<span style="color: turquoise ;">`r us_list[[7]][5,2]`</span>). 

## Contageous Spaces

#### What states have the **highest percent** of cases? What states have the **highest percent** of covid related deaths? What states have the **highest percent** of people fully and partially vaccinated? 

As of `r us_data_date`: 

`r us_list[[2]][1,1]` (<span style="color: turquoise ;">`r us_list[[2]][1,2]`</span>)  has the highest percent of cases per population followed by, `r us_list[[2]][2,1]` (<span style="color: turquoise ;">`r us_list[[2]][2,2]`</span>), `r us_list[[2]][3,1]` (<span style="color: turquoise ;">`r us_list[[2]][3,2]`</span>), `r us_list[[2]][4,1]` (<span style="color: turquoise ;">`r us_list[[2]][4,2]`</span>), and `r us_list[[2]][5,1]` (<span style="color: turquoise ;">`r us_list[[2]][5,2]`</span>). 

`r us_list[[4]][1,1]` (<span style="color: turquoise ;">`r us_list[[4]][1,2]`</span>) has the highest percent of covid related deaths per population followed by, `r us_list[[4]][2,1]` (<span style="color: turquoise ;">`r us_list[[4]][2,2]`</span>), `r us_list[[4]][3,1]` (<span style="color: turquoise ;">`r us_list[[4]][3,2]`</span>), `r us_list[[4]][4,1]` (<span style="color: turquoise ;">`r us_list[[4]][4,2]`</span>), and `r us_list[[4]][5,1]` (<span style="color: turquoise ;">`r us_list[[4]][5,2]`</span>). 

`r us_list[[6]][1,1]` (<span style="color: turquoise ;">`r us_list[[6]][1,2]`</span>) has the highest percent of people fully vaccinated, followed by `r us_list[[6]][2,1]` (<span style="color: turquoise ;">`r us_list[[6]][2,2]`</span>), `r us_list[[6]][3,1]` (<span style="color: turquoise ;">`r us_list[[6]][3,2]`</span>), `r us_list[[6]][4,1]` (<span style="color: turquoise ;">`r us_list[[6]][4,2]`</span>), and then `r us_list[[6]][5,1]` (<span style="color: turquoise ;">`r us_list[[6]][5,2]`</span>). 

`r us_list[[8]][1,1]` (<span style="color: turquoise ;">`r us_list[[8]][1,2]`</span>) has the highest percent of people partially vaccinated, followed by `r us_list[[8]][2,1]` (<span style="color: turquoise ;">`r us_list[[8]][2,2]`</span>), `r us_list[[8]][3,1]` (<span style="color: turquoise ;">`r us_list[[8]][3,2]`</span>), `r us_list[[8]][4,1]` (<span style="color: turquoise ;">`r us_list[[8]][4,2]`</span>), and then `r us_list[[8]][5,1]` (<span style="color: turquoise ;">`r us_list[[8]][5,2]`</span>). 

## Influx Dates

What days saw the highest number of new cases? What days saw the highes number of covid related deaths? What days saw highest number of people fully and partially vaccinated? 

[Coming Soon :D]

# Oregon

## Populated Places

#### What counties in Oreogn have the **most** covid cases? What counties in Oregon have the **most** deaths? 

`r or_list[[1]][1,1]` county (<span style="color: turquoise ;">`r or_list[[1]][1,2]`</span>) has had the most reported covid cases followed by `r or_list[[1]][2,1]` (<span style="color: turquoise ;">`r or_list[[1]][2,2]`</span>), `r or_list[[1]][3,1]` (<span style="color: turquoise ;">`r or_list[[1]][3,2]`</span>), `r or_list[[1]][4,1]` (<span style="color: turquoise ;">`r or_list[[1]][4,2]`</span>), and then `r or_list[[1]][5,1]` (<span style="color: turquoise ;">`r or_list[[1]][5,2]`</span>).

`r or_list[[3]][1,1]` county (<span style="color: turquoise ;">`r or_list[[3]][1,2]`</span>), has had the most most reported covid related deaths followed by `r or_list[[3]][2,1]` (<span style="color: turquoise ;">`r or_list[[3]][2,2]`</span>), `r or_list[[3]][3,1]` (<span style="color: turquoise ;">`r or_list[[3]][3,2]`</span>), `r or_list[[3]][4,1]` (<span style="color: turquoise ;">`r or_list[[3]][4,2]`</span>), and then `r or_list[[3]][5,1]` (<span style="color: turquoise ;">`r or_list[[3]][5,2]`</span>). 

## Contageous Spaces

#### What counties in Oregon have the **highest percent** of covid cases?What counties in Oregon have the **highest percent** of covid related deaths? 

As of `r or_data_date`:

`r or_list[[2]][1,1]` (<span style="color: turquoise ;">`r or_list[[2]][1,2]`</span>) county has the highest percent of cases per population followed by, `r or_list[[2]][2,1]` (<span style="color: turquoise ;">`r or_list[[2]][2,2]`</span>), `r or_list[[2]][3,1]` (<span style="color: turquoise ;">`r or_list[[2]][3,2]`</span>), `r or_list[[2]][4,1]` (<span style="color: turquoise ;">`r or_list[[2]][4,2]`</span>), and `r or_list[[2]][5,1]` (<span style="color: turquoise ;">`r or_list[[2]][5,2]`</span>). 

`r or_list[[4]][1,1]` (<span style="color: turquoise ;">`r or_list[[4]][1,2]`</span>) county has the highest percent of covid related deaths per population followed by `r or_list[[4]][2,1]` (<span style="color: turquoise ;">`r or_list[[4]][2,2]`</span>), `r or_list[[4]][3,1]` (<span style="color: turquoise ;">`r or_list[[4]][3,2]`</span>), `r or_list[[4]][4,1]` (<span style="color: turquoise ;">`r or_list[[4]][4,2]`</span>), and `r or_list[[4]][5,1]` (<span style="color: turquoise ;">`r or_list[[4]][5,2]`</span>). 

## Influx Dates

#### What days in Oregon saw the highest number of new cases? What days in Oregon saw the highest number of coivd related deaths? 

[Coming Soon :D]

