---
title: "About"
knit: (function(input_file, encoding) {
  out_dir <- '../docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
output:
  html_document:
    toc: false
    theme: darkly
---

```{r setup, include=FALSE}
here::i_am("04_website/index.Rmd")
```

```{r load data, include=FALSE}
# location of data
location_of_dates_data <- here::here("01_data/clean_data",
                                     "clean_data_dates.rds")
# save data
todays_date <- base::readRDS(location_of_dates_data)[[1]]
us_close_date <- base::readRDS(location_of_dates_data)[[2]]
or_close_date <- base::readRDS(location_of_dates_data)[[3]]
us_data_date <- base::readRDS(location_of_dates_data)[[4]]
or_data_date <- base::readRDS(location_of_dates_data)[[5]]
# data
todays_date
us_close_date
or_close_date
us_data_date
or_data_date
  # diff data 
us_date_diff <- todays_date - us_close_date
or_date_diff <- todays_date - or_close_date
```

The Coronavirus is an infectious disease^[**World Health Organization**  [_Coronavirus disease (COVID-19)_](https://www.who.int/health-topics/coronavirus#tab=tab_1). 2022.], that shut down the United States on `r us_close_date` (`r us_date_diff` days ago) and Oregon on `r or_close_date`(`r or_date_diff` days ago). While most of the state, country, and world have reopened, it appears that Covid might be sticking around longer than expected. This study uses two data sets from the NY Times github repository  ^[[New York Times Covid 19 Data](https://github.com/nytimes/covid-19-data).], and census data from the RStudio package tidycensus^[[tidycensus](https://walker-data.com/tidycensus/).], to create various data visualizations of covid in the united states, and Oregon. 

---

<center>

Project Updated: `r todays_date` $\quad$    |   $\quad$ US Data Updated: `r us_data_date`   $\quad$|$\quad$   County Data Update : `r or_data_date`

</center>
