---
title: "Analysis"
knit: (function(input_file, encoding) {
  out_dir <- '../docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'analysis.html'))})
output:
  html_document:
    toc: true
    toc_float: true
    theme: darkly
---

This analysis has two areas of focus: the United States, and Oregon counties. 

```{r, include=FALSE}
here::i_am("website/analysis.Rmd")
```

```{r load data, include=FALSE}
# data 
us_list <- readRDS(here::here("clean_data",
                              "us_list.rds"))
or_list <- readRDS(here::here("clean_data",
                              "or_list.rds"))
# variables
todays_date <- readRDS(here::here("clean_data",
                                  "date_today.rds"))
datas_date <- readRDS(here::here("clean_data",
                                 "date_data.rds"))
```

# United States

## Populated Places

#### What states have the **most** covid cases? What states have the **most** covid related deaths? What states have the **most** people fully and partially vaccinated? 

As of `r datas_date` the state with the most cases is `r us_list[[1]][1,1]` with `r us_list[[1]][1,2]` cases.  Followed by `r us_list[[1]][2,1]` with `r us_list[[1]][2,2]` cases, `r us_list[[1]][3,1]` with `r us_list[[1]][3,2]`, `r us_list[[1]][4,1]` with `r us_list[[1]][4,2]`, and then `r us_list[[1]][5,1]` with `r us_list[[1]][5,2]` number of cases. 


## Contageous Spaces

#### What states have the **highest percent** of cases? What states have the **highest percent** of covid related deaths? What states have the **highest percent** of people fully and partially vaccinated? 

As of `r datas_date` `r us_list[[2]][1,1]` has the highest percent of cases per population with `r us_list[[2]][1,2]`. Then `r us_list[[2]][2,1]` (`r us_list[[2]][2,2]`), `r us_list[[2]][3,1]` (`r us_list[[2]][3,2]`), `r us_list[[2]][4,1]` (`r us_list[[2]][4,2]`), and `r us_list[[2]][5,1]` (`r us_list[[2]][5,2]`). 


## Influx Dates

What days saw the highest number of new cases? What days saw the highes number of covid related deaths? What days saw highest number of people fully and partially vaccinated? 



# Oregon

## Populated Places

What counties in Oreogn have the **most** covid cases? What counties in Oregon have the **most** deaths? 


## Contageous Spaces

What counties in Oregon have the **highest percent** of covid cases?What counties in Oregon have the **highest percent** of covid related deaths? 


## Influx Dates

What days in Oregon saw the highest number of new cases? What days in Oregon saw the highest number of coivd related deaths? 








