---
title: "COVID-19"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
output:
  html_document:
    toc: true
    toc_float: TRUE
    theme: darkly
---

```{r, echo=F}
Date <- "2022-05-10"
N <- 5
```

```{r setup, include=FALSE}
here::i_am("website/index.Rmd")

knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(magrittr) # %>%
library(dplyr) # rename()
library(readr) # read_csv()
library(tidyr)
library(lubridate) # for dates
library(pander)
library(ggplot2)
library(gt) #format tables
library(kableExtra)
library(gridExtra) # grid.arrange
library(tidycensus) # population data
readRenviron(".Renviron")
library(plotly)
```


## Background and Data

COVID-19 is causing havoc in Oregon once again, and as numbers continue to spike, I decided to revisit one the of the first projects I worked on in R Studio. That project can be seen [here](https://rbolt13.github.io/rsite/covid.html), and used data from Johns Hopkins. However, because that data is no longer updated this investigation will use data from the NY times that has more current data. The repository for the NY Times data can be found [here](https://github.com/nytimes/covid-19-data), and the datasets that are being included are : 

1. **us.states** : state level data (file description [here](https://github.com/nytimes/covid-19-data/blob/master/README.md))

2. **us.counties** : county-level data (file description [here](https://github.com/nytimes/covid-19-data/blob/master/README.md))

3. **colleges** : number of reported cases among students and employees at American colleges and universities, updated May 26th (file description [here](https://github.com/nytimes/covid-19-data/tree/master/colleges))

4. **mask_use** : survey between July 2 and July 14 (2020) where participants were asked, _"How oftern do you wear a mask in public when you expect to be within six feet of another person?"_ (file description [here](https://github.com/nytimes/covid-19-data/tree/master/mask-use)) 

5. **vacc**: state level COVID-19 daily vaccination numbers time series data from the Johns Hopkins University repository (file description [here](https://github.com/govex/COVID-19/blob/master/data_tables/vaccine_data/us_data/readme.md))

6. **policytrackerOR**:  dates and description of policies going into/out of effect in Oregon.  To load data for a particular state go to [here](https://github.com/govex/COVID-19/tree/govex_data/data_tables/policy_data/table_data/Current), find the name of the state file you want to work with. 

Here is the data :

```{r load_data}
us.states <- read_csv('https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv')
us.counties <- read_csv('https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv')
colleges <- read_csv('https://raw.githubusercontent.com/nytimes/covid-19-data/master/colleges/colleges.csv')
mask_use <- read_csv('https://raw.githubusercontent.com/nytimes/covid-19-data/master/colleges/colleges.csv')
vacc <- read_csv("https://raw.githubusercontent.com/govex/COVID-19/master/data_tables/vaccine_data/us_data/time_series/people_vaccinated_us_timeline.csv")
policytrackerOR <- read_csv("https://raw.githubusercontent.com/govex/COVID-19/govex_data/data_tables/policy_data/table_data/Current/Oregon_policy.csv")
```

The projected cited above mainly looked at the case, death, and vaccination numbers per state to compare highly and mildly impacted states. In this project I will look at highly impacted states and the counties of Oregon. Additionally this project will use population and density data from the tidycensus package. I discuss more about how I got this data using an API in a blog post [here](https://rbolt.netlify.app/post/2021/08/30/tidycensus-package/). Note that this data is from 2019, which is a couple years older than the COVID data. 

```{r}
# tidycensus
# State : POP and DENSITY data 
state.pop <- get_estimates(geography = "state", year = 2019, variable =  "POP") %>% rename ("state" = NAME, "population" = value)
state.den <- get_estimates(geography = "state", year = 2019, variable =  "DENSITY") %>% rename ("state" = NAME, "density" = value)
# OREGON : POP and Density data 
or.county.pop <- get_estimates(geography = "county", state = "OR", year = 2019, variable = "POP") %>% rename ("county" = NAME, "population" = value)
or.county.den <- get_estimates(geography = "county", state = "OR", year = 2019, variable = "DENSITY") %>% rename ("county" = NAME, "density" = value)
```

# Wrangling the Data
The COVID data set is already in long form (meaning the dates are in rows instead of columns), and the date is already saved as a date variable. Therefore the main tasks here are to join the us.states data set with the vaccination records, and population estimates. Then join the us.states.vacc with the population data and create new percentage columns. 

```{r, echo=F}
# join states population with density
state.est <- state.pop %>%
  select(state, population) %>%
  full_join(state.den %>%
              select(state, density),
            by = c("state" = "state"))
# join states with vacc
us.states.vacc <- us.states %>%
  select(date, state, cases, deaths) %>%
  full_join(vacc %>%
              select(Province_State, Date, Lat, Long_,People_Fully_Vaccinated, People_Partially_Vaccinated),
            by = c("state"="Province_State", "date"="Date")) %>%
  select(date, state, cases, deaths, People_Fully_Vaccinated, People_Partially_Vaccinated) %>%
  # join with census population data
  full_join(state.est %>%
              select(state, population, density),
            by = c("state"="state")) %>%
  summarise(date = date,
            state = state,
            population = population,
            density = density,
            cases = cases,
            case.per = cases/population,
            deaths = deaths,
            perc.deaths = deaths/population,
            full.vacc = People_Fully_Vaccinated,
            full.vacc.perc = full.vacc/population,
            part.vacc = People_Partially_Vaccinated,
            part.vacc.perc = part.vacc/population)
head(us.states.vacc, n=5)
```

Note: People_Fully_Vaccinated and People_Partially_Vaccinated show as NA because of the dates displayed were before the vaccine was released.

Next, to join the Oregon county population data with the density data, remove "County, Oregon" from each county object, filter out Oregon from the us.counties data, and then join with the population data per county. 

```{r, echo=FALSE}
# join states population with density and clean
or.county.est <- or.county.pop %>%
  select(county, population) %>%
  full_join(or.county.den %>%
              select(county, density),
            by = c("county" = "county"))
# remove " County, Oregon" from the county name (this will make joining later easier)
or.county.est$county <- gsub(" County, Oregon", "", or.county.pop$county )
# select only oregon counties, and join with population
or.covid.est <- us.counties %>%
  filter(state == "Oregon") %>%
  select(date, county, cases, deaths) %>%
  full_join(or.county.est %>%
              select(county, population, density),
            by = c("county"="county"))%>% 
  summarise(date = date,
            county = county,
            population = population,
            density = density,
            cases = cases,
            cases.perc = cases/population,
            deaths = deaths,
            deaths.perc = deaths/population) 
head(or.covid.est, n=10)
```

---

# Using the data

## Looking at States

To begin lets look at the country as a whole, by state. The data will be filtered for **`r Date`**, and then lets look at the top five states with :

### Most number of cases

```{r, echo=F}
us.states.cases <- us.states.vacc %>%
  filter(date == Date) %>%
  select(state, population, cases, case.per) %>%
  arrange(desc(cases)) %>%
    head(5) %>%
    gt() %>%
    fmt_number(columns = cases, decimals = 0) %>%
    fmt_number(columns = population, decimals = 0) %>%
    fmt_percent(columns = case.per, decimals = 2) %>%
    tab_header(title = md("**Most Cases**"),
               subtitle = paste0("By State as of ", Date)) %>%
    cols_label(state = md("**State**"),
               population = md("**Total Population**"),
               cases = md("**Cases**"),
               case.per = md("**Percentage**")) %>%
   tab_style(style = list(
             cell_fill(color = "lightcyan")),
   locations = cells_body(
       columns = everything()
     ))
us.states.cases
```

### Highest percentage of cases 
```{r, echo=F}
# filter - select - arrange
us.states.cases.perc <- us.states.vacc %>%
  filter(date == Date) %>%
  select(state, population, cases, case.per) %>%
  arrange(desc(case.per)) 
# table
gt(us.states.cases.perc[1:5,]) %>%
    fmt_number(columns = cases, decimals = 0) %>%
    fmt_number(columns = population, decimals = 0) %>%
    fmt_percent(columns = case.per, decimals = 2) %>%
    tab_header(title = md("**Highest Percent of Cases**"),
               subtitle = paste0("By State as of ", Date)) %>%
    cols_label(state = md("**State**"),
               population = md("**Total Population**"),
               cases = md("**Cases**"),
               case.per = md("**Percentage**")) %>%
   tab_style(style = list(
             cell_fill(color = "lightcyan")),
   locations = cells_body(
       columns = everything()
     ))
```

### Most number of deaths
```{r, echo=F}
us.states.deaths <- us.states.vacc %>%
  filter(date == Date) %>%
  select(state, population, deaths, perc.deaths) %>%
  arrange(desc(deaths))
# table
gt(us.states.deaths[1:5,]) %>%
    fmt_number(columns = deaths, decimals = 0) %>%
    fmt_number(columns = population, decimals = 0) %>%
    fmt_percent(columns = perc.deaths, decimals = 2) %>%
    tab_header(title = md("**Most Deaths**"),
               subtitle = paste0("By State as of ", Date)) %>%
    cols_label(state = md("**State**"),
               population = md("**Total Population**"),
               deaths = md("**Deaths**"),
               perc.deaths = md("**Percentage**")) %>%
   tab_style(style = list(
             cell_fill(color = "lightcyan")),
   locations = cells_body(
       columns = everything()
     ))
```

### Highest percentage of deaths 
```{r, echo=F}
us.states.deaths.perc <- us.states.vacc %>%
  filter(date == Date) %>%
  select(state, population, deaths, perc.deaths) %>%
  arrange(desc(perc.deaths)) 
# table
gt(us.states.deaths.perc[1:5,]) %>%
    fmt_number(columns = deaths, decimals = 0) %>%
    fmt_number(columns = population, decimals = 0) %>%
    fmt_percent(columns = perc.deaths, decimals = 2) %>%
    tab_header(title = md("**Highest percent of Deaths**"),
               subtitle = paste0("By State as of ", Date)) %>%
    cols_label(state = md("**State**"),
               population = md("**Total Population**"),
               deaths = md("**Deaths**"),
               perc.deaths = md("**Percentage**")) %>%
   tab_style(style = list(
             cell_fill(color = "lightcyan")),
   locations = cells_body(
       columns = everything()
     ))
```

### Most people fully vaccinated 
```{r, echo=F}
us.states.most.vacc <- us.states.vacc %>%
  filter(date == Date) %>%
  select(state, population, full.vacc, full.vacc.perc) %>%
  arrange(desc(full.vacc)) 
# Table
gt(us.states.most.vacc[1:5,]) %>%
    fmt_number(columns = full.vacc, decimals = 0) %>%
    fmt_number(columns = population, decimals = 0) %>%
    fmt_percent(columns = full.vacc.perc, decimals = 2) %>%
    tab_header(title = md("**Most People Vaccinated**"),
               subtitle = paste0("By State as of ", Date)) %>%
    cols_label(state = md("**State**"),
               population = md("**Total Population**"),
               full.vacc = md("**People Fully Vaccinated**"),
               full.vacc.perc = md("**Percentage**")) %>%
   tab_style(style = list(
             cell_fill(color = "lightcyan")),
   locations = cells_body(
       columns = everything()
     ))
```

### Highest percentage of population fully vaccinated 

```{r, echo=F}
us.states.vacc.perc <- us.states.vacc %>%
  filter(date == Date) %>%
  select(state, population, full.vacc, full.vacc.perc) %>%
  arrange(desc(full.vacc.perc))
# table
gt(us.states.vacc.perc[1:5,]) %>%
    fmt_number(columns = full.vacc, decimals = 0) %>%
    fmt_number(columns = population, decimals = 0) %>%
    fmt_percent(columns = full.vacc.perc, decimals = 2) %>%
    tab_header(title = md("**Highest Percent of People Fully Vaccinated**"),
               subtitle = paste0("By State as of ", Date)) %>%
    cols_label(state = md("**State**"),
               population = md("**Total Population**"),
               full.vacc = md("**People Fully Vaccinated**"),
               full.vacc.perc = md("**Percentage**")) %>%
   tab_style(style = list(
             cell_fill(color = "lightcyan")),
   locations = cells_body(
       columns = everything()
     ))
```

## Looking at Oregon Counties

```{r data for maps, echo = F , message= F}
# remove " County, Oregon" from the county name
or.county.pop$county <- gsub(" County, Oregon", "", or.county.pop$county)
# sort or.county.pop
# or.county.pop <- or.county.pop[with(or.county.pop, order(county)), ]
# or counties with lat and long
or_counties <- map_data("county", "oregon") %>%
  select(lon = long, lat, group, county = subregion)
# total cases and deaths 
total_or <- or.covid.est %>%
  filter(date == Date) %>%
  group_by(county) %>%
  summarise(total_cases = cases,
            total_cases_perc = cases.perc,
            total_deaths = deaths,
            total_deaths_perc = deaths.perc)
# remove "unknown" from total cases 
total_or <- total_or %>%
  filter(county != "Unknown")
# join total_or with population
total_or <- total_or %>%
  select(county, total_cases, total_cases_perc, total_deaths, total_deaths_perc) %>%
  full_join(or.county.pop %>%
              select(county, population),
            by = c("county"="county"))

```


```{r, echo=F, message=F}
# join or_counties with total_or
new <- or_counties %>%
  select(lon, lat, group, county) %>%
  left_join(total_or %>%
            select(county, total_cases, total_cases_perc, total_deaths, total_deaths_perc),
            by = c("county" = "county"))
```


```{r, echo=F, message=F}
# baker data 
baker_tcase <- total_or$total_cases[1]
baker_tcasep <- total_or$total_cases_perc[1]
baker_tdeath <- total_or$total_deaths[1]
baker_deathp <- total_or$total_deaths_perc[1]
new$total_cases[1:73] <- baker_tcase
new$total_cases_perc[1:73] <- baker_tcasep
new$total_deaths[1:73] <- baker_tdeath
new$total_deaths_perc[1:73] <- baker_deathp
# benton
benton_tcase <- total_or$total_cases[2]
benton_tcasep <- total_or$total_cases_perc[2]
benton_tdeath <- total_or$total_deaths[2]
benton_deathp <- total_or$total_deaths_perc[2]
new$total_cases[74:106] <- benton_tcase
new$total_cases_perc[74:106] <- benton_tcasep
new$total_deaths[74:106] <- benton_tdeath
new$total_deaths_perc[74:106] <- benton_deathp
# clackamas
clackamas_tcase <- total_or$total_cases[3]
clackamas_tcasep <- total_or$total_cases_perc[3]
clackamas_tdeath <- total_or$total_deaths[3]
clackamas_deathp <- total_or$total_deaths_perc[3]
new$total_cases[107:156] <- clackamas_tcase
new$total_cases_perc[107:156] <- clackamas_tcasep
new$total_deaths[107:156] <- clackamas_tdeath
new$total_deaths_perc[107:156] <- clackamas_deathp
# clatsop
clatsop_tcase <- total_or$total_cases[4]
clatsop_tcasep <- total_or$total_cases_perc[4]
clatsop_tdeath <- total_or$total_deaths[4]
clatsop_deathp <- total_or$total_deaths_perc[4]
new$total_cases[157:195] <- clatsop_tcase
new$total_cases_perc[157:195] <- clatsop_tcasep
new$total_deaths[157:195] <- clatsop_tdeath
new$total_deaths_perc[157:195] <- clatsop_deathp
# columbia
columbia_tcase <- total_or$total_cases[5]
columbia_tcasep <- total_or$total_cases_perc[5]
columbia_tdeath <- total_or$total_deaths[5]
columbia_deathp <- total_or$total_deaths_perc[5]
new$total_cases[196:225] <- columbia_tcase
new$total_cases_perc[196:225] <- columbia_tcasep
new$total_deaths[196:225] <- columbia_tdeath
new$total_deaths_perc[196:225] <- columbia_deathp
# coos - 6, 226:271
coos_tcase <- total_or$total_cases[6]
coos_tcasep <- total_or$total_cases_perc[6]
coos_tdeath <- total_or$total_deaths[6]
coos_deathp <- total_or$total_deaths_perc[6]
new$total_cases[226:271] <- coos_tcase
new$total_cases_perc[226:271] <- coos_tcasep
new$total_deaths[226:271] <- coos_tdeath
new$total_deaths_perc[226:271] <- coos_deathp
# crook - 7, 272:300
crook_tcase <- total_or$total_cases[7]
crook_tcasep <- total_or$total_cases_perc[7]
crook_tdeath <- total_or$total_deaths[7]
crook_deathp <- total_or$total_deaths_perc[7]
new$total_cases[272:300] <- crook_tcase
new$total_cases_perc[272:300] <- crook_tcasep
new$total_deaths[272:300] <- crook_tdeath
new$total_deaths_perc[272:300] <- crook_deathp
# curry - 8, 301:368
curry_tcase <- total_or$total_cases[8]
curry_tcasep <- total_or$total_cases_perc[8]
curry_tdeath <- total_or$total_deaths[8]
curry_deathp <- total_or$total_deaths_perc[8]
new$total_cases[301:368] <- curry_tcase
new$total_cases_perc[301:368] <- curry_tcasep
new$total_deaths[301:368] <- curry_tdeath
new$total_deaths_perc[301:368] <- curry_deathp
# deschutes - 9, 369:418
deschutes_tcase <- total_or$total_cases[9]
deschutes_tcasep <- total_or$total_cases_perc[9]
deschutes_tdeath <- total_or$total_deaths[9]
deschutes_deathp <- total_or$total_deaths_perc[9]
new$total_cases[369:418] <- deschutes_tcase
new$total_cases_perc[369:418] <- deschutes_tcasep
new$total_deaths[369:418] <- deschutes_tdeath
new$total_deaths_perc[369:418] <- deschutes_deathp
# douglas - 10, 419:515
douglas_tcase <- total_or$total_cases[10]
douglas_tcasep <- total_or$total_cases_perc[10]
douglas_tdeath <- total_or$total_deaths[10]
douglas_deathp <- total_or$total_deaths_perc[10]
new$total_cases[419:515] <- douglas_tcase
new$total_cases_perc[419:515] <- douglas_tcasep
new$total_deaths[419:515] <- douglas_tdeath
new$total_deaths_perc[419:515] <- douglas_deathp
# gilliam - 11, 516:566
gilliam_tcase <- total_or$total_cases[11]
gilliam_tcasep <- total_or$total_cases_perc[11]
gilliam_tdeath <- total_or$total_deaths[11]
gilliam_deathp <- total_or$total_deaths_perc[11]
new$total_cases[516:566] <- gilliam_tcase
new$total_cases_perc[516:566] <- gilliam_tcasep
new$total_deaths[516:566] <- gilliam_tdeath
new$total_deaths_perc[516:566] <- gilliam_deathp
# grant - 12, 567:616
grant_tcase <- total_or$total_cases[12]
grant_tcasep <- total_or$total_cases_perc[12]
grant_tdeath <- total_or$total_deaths[12]
grant_deathp <- total_or$total_deaths_perc[12]
new$total_cases[567:616] <- grant_tcase
new$total_cases_perc[567:616] <- grant_tcasep
new$total_deaths[567:616] <- grant_tdeath
new$total_deaths_perc[567:616] <- grant_deathp
# harney - 13, 617:645
harney_tcase <- total_or$total_cases[13]
harney_tcasep <- total_or$total_cases_perc[13]
harney_tdeath <- total_or$total_deaths[13]
harney_deathp <- total_or$total_deaths_perc[13]
new$total_cases[617:645] <- harney_tcase
new$total_cases_perc[617:645] <- harney_tcasep
new$total_deaths[617:645] <- harney_tdeath
new$total_deaths_perc[617:645] <- harney_deathp
# hoodriver - 14, 646:674
hoodriver_tcase <- total_or$total_cases[14]
hoodriver_tcasep <- total_or$total_cases_perc[14]
hoodriver_tdeath <- total_or$total_deaths[14]
hoodriver_deathp <- total_or$total_deaths_perc[14]
new$total_cases[646:674] <- hoodriver_tcase
new$total_cases_perc[646:674] <- hoodriver_tcasep
new$total_deaths[646:674] <- hoodriver_tdeath
new$total_deaths_perc[646:674] <- hoodriver_deathp
# jackson - 15, 675:694
a <- total_or$total_cases[15]
b <- total_or$total_cases_perc[15]
c <- total_or$total_deaths[15]
d <- total_or$total_deaths_perc[15]
new$total_cases[675:694] <- a
new$total_cases_perc[675:694] <- b
new$total_deaths[675:694] <- c
new$total_deaths_perc[675:694] <- d
# jefferson - 16, 695:714
a <- total_or$total_cases[16]
b <- total_or$total_cases_perc[16]
c <- total_or$total_deaths[16]
d <- total_or$total_deaths_perc[16]
new$total_cases[695:714] <- a
new$total_cases_perc[695:714] <- b
new$total_deaths[695:714] <- c
new$total_deaths_perc[695:714] <- d
```

```{r, echo=F, message= F}
# josephine - 17, 715:758
a <- total_or$total_cases[17]
b <- total_or$total_cases_perc[17]
c <- total_or$total_deaths[17]
d <- total_or$total_deaths_perc[17]
new$total_cases[715:758] <- a
new$total_cases_perc[715:758] <- b
new$total_deaths[715:758] <- c
new$total_deaths_perc[715:758] <- d
# klamath - 18, 759:795
a <- total_or$total_cases[18]
b <- total_or$total_cases_perc[18]
c <- total_or$total_deaths[18]
d <- total_or$total_deaths_perc[18]
new$total_cases[759:795] <- a
new$total_cases_perc[759:795] <- b
new$total_deaths[759:795] <- c
new$total_deaths_perc[759:795] <- d
# lane - 19, 796:818
a <- total_or$total_cases[19]
b <- total_or$total_cases_perc[19]
c <- total_or$total_deaths[19]
d <- total_or$total_deaths_perc[19]
new$total_cases[796:818] <- a
new$total_cases_perc[796:818] <- b
new$total_deaths[796:818] <- c
new$total_deaths_perc[796:818] <- d
# lake - 20, 819:908
a <- total_or$total_cases[20]
b <- total_or$total_cases_perc[20]
c <- total_or$total_deaths[20]
d <- total_or$total_deaths_perc[20]
new$total_cases[819:908] <- a
new$total_cases_perc[819:908] <- b
new$total_deaths[819:908] <- c
new$total_deaths_perc[819:908] <- d
# lincoln - 21, 909:987
a <- total_or$total_cases[21]
b <- total_or$total_cases_perc[21]
c <- total_or$total_deaths[21]
d <- total_or$total_deaths_perc[21]
new$total_cases[909:987] <- a
new$total_cases_perc[909:987] <- b
new$total_deaths[909:987] <- c
new$total_deaths_perc[909:987] <- d
# linn - 22, 988:1020
a <- total_or$total_cases[22]
b <- total_or$total_cases_perc[22]
c <- total_or$total_deaths[22]
d <- total_or$total_deaths_perc[22]
new$total_cases[988:1020] <- a
new$total_cases_perc[988:1020] <- b
new$total_deaths[988:1020] <- c
new$total_deaths_perc[988:1020] <- d
# malheur - 23, 1021:1058
a <- total_or$total_cases[23]
b <- total_or$total_cases_perc[23]
c <- total_or$total_deaths[23]
d <- total_or$total_deaths_perc[23]
new$total_cases[1021:1058] <- a
new$total_cases_perc[1021:1058] <- b
new$total_deaths[1021:1058] <- c
new$total_deaths_perc[1021:1058] <- d
# marion - 24, 1059:1207
a <- total_or$total_cases[24]
b <- total_or$total_cases_perc[24]
c <- total_or$total_deaths[24]
d <- total_or$total_deaths_perc[24]
new$total_cases[1059:1207] <- a
new$total_cases_perc[1059:1207] <- b
new$total_deaths[1059:1207] <- c
new$total_deaths_perc[1059:1207] <- d
# marrow - 25, 1127:1155
a <- total_or$total_cases[25]
b <- total_or$total_cases_perc[25]
c <- total_or$total_deaths[25]
d <- total_or$total_deaths_perc[25]
new$total_cases[1127:1155] <- a
new$total_cases_perc[1127:1155] <- b
new$total_deaths[1127:1155] <- c
new$total_deaths_perc[1127:1155] <- d
# multnomah - 26, 1156:1181
a <- total_or$total_cases[26]
b <- total_or$total_cases_perc[26]
c <- total_or$total_deaths[26]
d <- total_or$total_deaths_perc[26]
new$total_cases[1156:1181] <- a
new$total_cases_perc[1156:1181] <- b
new$total_deaths[1156:1181] <- c
new$total_deaths_perc[1156:1181] <- d
# polk - 27, 1182:1207
a <- total_or$total_cases[27]
b <- total_or$total_cases_perc[27]
c <- total_or$total_deaths[27]
d <- total_or$total_deaths_perc[27]
new$total_cases[1182:1207] <- a
new$total_cases_perc[1182:1207] <- b
new$total_deaths[1182:1207] <- c
new$total_deaths_perc[1182:1207] <- d
# sherman - 28, 1208:1258
a <- total_or$total_cases[28]
b <- total_or$total_cases_perc[28]
c <- total_or$total_deaths[28]
d <- total_or$total_deaths_perc[28]
new$total_cases[1208:1258] <- a
new$total_cases_perc[1208:1258] <- b
new$total_deaths[1208:1258] <- c
new$total_deaths_perc[1208:1258] <- d
# tillamook - 29, 1259:1308
a <- total_or$total_cases[29]
b <- total_or$total_cases_perc[29]
c <- total_or$total_deaths[29]
d <- total_or$total_deaths_perc[29]
new$total_cases[1259:1308] <- a
new$total_cases_perc[1259:1308] <- b
new$total_deaths[1259:1308] <- c
new$total_deaths_perc[1259:1308] <- d
# umatilla - 30, 1309:1351
a <- total_or$total_cases[30]
b <- total_or$total_cases_perc[30]
c <- total_or$total_deaths[30]
d <- total_or$total_deaths_perc[30]
new$total_cases[1309:1351] <- a
new$total_cases_perc[1309:1351] <- b
new$total_deaths[1309:1351] <- c
new$total_deaths_perc[1309:1351] <- d
# union - 31, 1352:1416
a <- total_or$total_cases[31]
b <- total_or$total_cases_perc[31]
c <- total_or$total_deaths[31]
d <- total_or$total_deaths_perc[31]
new$total_cases[1352:1416] <- a
new$total_cases_perc[1352:1416] <- b
new$total_deaths[1352:1416] <- c
new$total_deaths_perc[1352:1416] <- d
# wallowa - 32, 1417:1465
a <- total_or$total_cases[32]
b <- total_or$total_cases_perc[32]
c <- total_or$total_deaths[32]
d <- total_or$total_deaths_perc[32]
new$total_cases[1417:1465] <- a
new$total_cases_perc[1417:1465] <- b
new$total_deaths[1417:1465] <- c
new$total_deaths_perc[1417:1465] <- d
# wasco - 33, 1466:1528
a <- total_or$total_cases[33]
b <- total_or$total_cases_perc[33]
c <- total_or$total_deaths[33]
d <- total_or$total_deaths_perc[33]
new$total_cases[1466:1528] <- a
new$total_cases_perc[1466:1528] <- b
new$total_deaths[1466:1528] <- c
new$total_deaths_perc[1466:1528] <- d
# washington - 34, 1529:1576
a <- total_or$total_cases[34]
b <- total_or$total_cases_perc[34]
c <- total_or$total_deaths[34]
d <- total_or$total_deaths_perc[34]
new$total_cases[1529:1576] <- a
new$total_cases_perc[1529:1576] <- b
new$total_deaths[1529:1576] <- c
new$total_deaths_perc[1529:1576] <- d
# wheeler - 35, 1577:1605
a <- total_or$total_cases[35]
b <- total_or$total_cases_perc[35]
c <- total_or$total_deaths[35]
d <- total_or$total_deaths_perc[35]
new$total_cases[1577:1605] <- a
new$total_cases_perc[1577:1605] <- b
new$total_deaths[1577:1605] <- c
new$total_deaths_perc[1577:1605] <- d
# yamhill - 36, 1606:1633
a <- total_or$total_cases[36]
b <- total_or$total_cases_perc[36]
c <- total_or$total_deaths[36]
d <- total_or$total_deaths_perc[36]
new$total_cases[1606:1633] <- a
new$total_cases_perc[1606:1633] <- b
new$total_deaths[1606:1633] <- c
new$total_deaths_perc[1606:1633] <- d
```



Next to look at the data a little closer to home, for Oregon Counties. Initially filtering by the most recent date, which as of this being written is **`r Date``**, looking at a graph of the state as a whole, and then look at the top five Oregon counties.

### Most number of cases

```{r, echo=F, message= F}
map <- ggplot(data = new, aes(lon, lat, group = group, fill = total_cases))+
  geom_polygon(color = "grey") +
  coord_quickmap() +
  theme_minimal() +
  ggtitle(paste("Total Number of Cases by Oregon Counties as of", Date)) +
  labs(x = "",
       y = "",
       fill = "Number of Cases") 
  
ggplotly(map) 
```

```{r, echo = F , message= F}
plot <- or.covid.est %>%
  ggplot(aes(x=date, y=cases, color = county)) +
  geom_line(size=.8) +
  ggtitle("Number of Cases per County") +
  xlab("Date (by month)") + ylab("Cases") +
  scale_x_date(date_breaks = "1 month", date_labels = "%m") +
  theme_minimal()
ggplotly(plot)
```

```{r, echo = F}
# cases
or.county.cases <- or.covid.est %>%
  filter(date == Date) %>%
  select(county, population, cases, cases.perc) %>%
  arrange(desc(cases)) %>%
    head(5) %>%
    gt() %>%
    fmt_number(columns = cases, decimals = 0) %>%
    fmt_number(columns = population, decimals = 0) %>%
    fmt_percent(columns = cases.perc, decimals = 2) %>%
    tab_header(title = md("**Most Cases**"),
               subtitle = paste0("Oregon Counties as of ", Date)) %>%
    cols_label(county = md("**County**"),
               population = md("**Total Population**"),
               cases = md("**Cases**"),
               cases.perc = md("**Percentage**")) %>%
   tab_style(style = list(
             cell_fill(color = "#F9E3D6")),
   locations = cells_body(
       columns = everything()
     ))
or.county.cases
```


### Highest percentage of Cases 

```{r, echo=F, message= F}
map <- ggplot(data = new, aes(lon, lat, group = group, fill = total_cases_perc))+
  geom_polygon(color = "grey") +
  coord_quickmap() +
  theme_minimal() +
  ggtitle(paste("Total % of Cases Per Oregon County as of", Date)) +
  labs(x = "",
       y = "",
       fill = "Percent of Cases")
ggplotly(map)
```

```{r, echo = F , message=F, warning=F}
plot2 <- or.covid.est %>%
  ggplot(aes(x=date, y=cases.perc, color = county)) +
  geom_line(size=.8) +
  ggtitle("Percent of Cases per County") +
  xlab("Date (by month)") + ylab("Percent") +
  scale_x_date(date_breaks = "1 month", date_labels = "%m") +
  theme_minimal()
ggplotly(plot2)
```

```{r, echo = F}
#cases percent
or.county.cases.perc <- or.covid.est %>%
  filter(date == Date) %>%
  select(county, population, cases, cases.perc) %>%
  arrange(desc(cases.perc)) %>%
    head(5) %>%
    gt() %>%
    fmt_number(columns = cases, decimals = 0) %>%
    fmt_number(columns = population, decimals = 0) %>%
    fmt_percent(columns = cases.perc, decimals = 2) %>%
    tab_header(title = md("**Highest Percentage of Cases**"),
               subtitle = paste0("Oregon Counties as of ", Date)) %>%
    cols_label(county = md("**County**"),
               population = md("**Total Population**"),
               cases = md("**Cases**"),
               cases.perc = md("**Percentage**")) %>%
   tab_style(style = list(
             cell_fill(color = "#F9E3D6")),
   locations = cells_body(
       columns = everything()
     ))
or.county.cases.perc
```

### Most number of Deaths

```{r, echo=F, message= F}
map <- ggplot(data = new, aes(lon, lat, group = group, fill = total_deaths))+
  geom_polygon(color = "grey") +
  coord_quickmap() +
  theme_minimal() +
  ggtitle(paste("Total Number of Deaths Per Oregon County as of", Date)) +
  labs(x = "",
       y = "",
       fill = "Number of Deaths")
ggplotly(map)
```

```{r, echo = F , message= F}
plot3 <- or.covid.est %>%
  ggplot(aes(x=date, y=deaths, color = county)) +
  geom_line(size=.8) +
  ggtitle("Number of Deaths per County") +
  xlab("Date (by month)") + ylab("Deaths") +
  scale_x_date(date_breaks = "1 month", date_labels = "%m") +
  theme_minimal()
ggplotly(plot3)
```

```{r, echo = F}
#deaths
or.county.deaths <- or.covid.est %>%
  filter(date == Date) %>%
  select(county, population, deaths, deaths.perc) %>%
  arrange(desc(deaths)) %>%
    head(5) %>%
    gt() %>%
    fmt_number(columns = deaths, decimals = 0) %>%
    fmt_number(columns = population, decimals = 0) %>%
    fmt_percent(columns = deaths.perc, decimals = 2) %>%
    tab_header(title = md("**Most Deaths**"),
               subtitle = paste0("Oregon Counties as of ", Date)) %>%
    cols_label(county = md("**County**"),
               population = md("**Total Population**"),
               deaths = md("**Deaths**"),
               deaths.perc = md("**Percentage**")) %>%
   tab_style(style = list(
             cell_fill(color = "#F9E3D6")),
   locations = cells_body(
       columns = everything()
     ))
or.county.deaths
```

### Highest percentage of Deaths

```{r, echo=F, message= F}
map <- ggplot(data = new, aes(lon, lat, group = group, fill = total_deaths_perc))+
  geom_polygon(color = "grey") +
  coord_quickmap() +
  theme_minimal() +
  ggtitle(paste("Total % of Deaths Per Oregon County as of", Date)) +
  labs(x = "",
       y = "",
       fill = "Percent of Deaths")
ggplotly(map)
```

```{r, echo = F , message= FALSE, error=F, warning=FALSE}
plot4 <- or.covid.est %>%
  ggplot(aes(x=date, y=deaths.perc, color = county)) +
  geom_line(size=.8) +
  ggtitle("Percent of Deaths per County") +
  xlab("Date (by month)") + ylab("Percent") +
  scale_x_date(date_breaks = "1 month", date_labels = "%m") +
  theme_minimal()
ggplotly(plot4)
```

```{r, echo = F}
#deaths percentage
or.county.deaths.perc <- or.covid.est %>%
  filter(date == Date) %>%
  select(county, population, deaths, deaths.perc) %>%
  arrange(desc(deaths.perc)) %>%
    head(5) %>%
    gt() %>%
    fmt_number(columns = deaths, decimals = 0) %>%
    fmt_number(columns = population, decimals = 0) %>%
    fmt_percent(columns = deaths.perc, decimals = 2) %>%
    tab_header(title = md("**Highest Percent of Deaths**"),
               subtitle = paste0("Oregon Counties as of ", Date)) %>%
    cols_label(county = md("**County**"),
               population = md("**Total Population**"),
               deaths = md("**Deaths**"),
               deaths.perc = md("**Percentage**")) %>%
   tab_style(
     style = list(
       cell_fill(color = "#F9E3D6")
       ),
     locations = cells_body(
       columns = everything()
     ))
or.county.deaths.perc
```

